Epoch of Reionization Power Spectrum Results from PAPER-128

### 2016-07-14 PAPER.tex
some paragraphs on delay filtering

Calibration of antennas in a redundant array is therefore crucial to maximize sensitivity. In our raw data, visibility measurements from two identical baseline types are not necessarily be equal. This is due to changing system temperatures and different signal chains as data flows from each antenna into the correlator. However, PAPER's redundant layout can be used to solve for complex antenna gains, which when applied to the data, bring all visibility measurements for a specific baseline type into alignment.

@@

Here $V_{ij}'$ represents a measured visibility for a baseline with antennas $i,j$, $V_{ij}$ represents the `true' sky visibility for that same baseline, $g_{i}$ and $g_{j}$ are the complex gain factors for each antenna, and $n_{ij}$ is noise. Each gain number consist of an amplitude $\eta$ and phase $\phi$. The visibility equation can be written down for all baselines and - without using array redundancy - would not be least squares solvable because of too many unknowns. However, if $V_{ij}$, the `true', or `model' visibility, is constrained to be the same for all identical baselines, the system of equations becomes least squares solvable.

@@

For PAPER-128, we update the Omnical software package, originally written in C, with a user-friendly interface written in Python. We also include the subtraction of crosstalk in our dataset. Crosstalk shows up as a constant signal in time in our visibilities and is attributed to signals from one antenna being picked up by another. Omnical handles crosstalk by subtracting off an averaged residual (differences between measured visibilities and the `model' visibility, averaged over baseline type) for each visibility measurement. Additionally, we employ a chi-square metric on a per antenna basis in order to exclude dead or faulty antennas and improve our calibration solutions. \cc{Add plots to show stability of Omnical solutions and chi-square examples}

We perform redundant calibration for each epoch of data used in our analysis, as each epoch has to be separately calibrated since they are on different overall gain scales. We divide out the antenna gains for each of our measured visibilities, bringing them into alignment with each other (for identical baseline types) and the `model' visibility for that baseline type. It is important to note that while Omnical is powerful for ensuring array redundancy, it is not able to solve for $4$ calibration parameters - namely, to correct for the overall gain, phase, and tip/tilt of the array. In order to do this, we turn to...

@@

>>>>

Calibration of antennas in a redundant array is therefore crucial in order to maximize sensitivity. In our raw data, visibility measurements from two identical baseline types are not necessarily equal. This is due to changing system temperatures and different signal chains as data flows from each antenna into the correlator. However, PAPER's redundant layout can be used to solve for complex antenna gains, which when applied to the data, bring all visibility measurements for a specific baseline type into alignment.

@@

Here $V_{ij}'$ represents a measured visibility for a baseline with antennas $i$ and $j$, $V_{ij}$ represents the `true' sky visibility for that same baseline, $g_{i}$ and $g_{j}$ are the complex gain factors for each antenna, and $n_{ij}$ is noise. Each gain number consists of an amplitude $\eta$ and phase $\phi$. The visibility equation can be written down for all baselines and - without using array redundancy - would not be least squares solvable because the number of unknowns ($2N+N(N-1)/2$) is greater than the number of measurements ($N(N-1)/2$), where $N$ is the number of antennas . However, if $V_{ij}$, the `true', or `model' visibility, is constrained to be the same for all identical baselines, the system of equations becomes solvable.

@@

We perform redundant calibration for each epoch of data used in our analysis, as each epoch has to be separately calibrated since they are on different overall gain scales. We divide out the antenna gains for each of our measured visibilities, bringing them into alignment with each other (for identical baseline types) and the `model' visibility for that baseline type. We also include the subtraction of crosstalk in our dataset. Crosstalk shows up as a constant signal in time in our visibilities and is attributed to signals from one antenna being picked up by another. Omnical handles crosstalk by subtracting off an averaged residual (differences between measured visibilities and the `model' visibility, averaged over baseline type) for each visibility measurement.

For PAPER-128, we update the Omnical software package, originally written in C, with a user-friendly interface written in Python.  Additionally, we employ a chi-square metric on a per antenna basis in order to exclude dead or faulty antennas and improve our calibration solutions. \cc{Add plots to show stability of Omnical solutions and chi-square examples}

It is important to note that while Omnical is powerful for ensuring array redundancy, it is not able to solve for $4$ calibration parameters - namely, to correct for the overall gain, phase, and tip/tilt of the array. In order to do this, we turn to...

@@

Foregrounds pose a significant challenge to EoR experiments. Some experiments, like LOFAR and the MWA, rely on modeling and foreground subtraction to create a clean EoR `window' in which to make their power spectrum measurements. Previous PAPER results, including this one, instead focus on a foreground avoidance approach.

In order to filter foregrounds, we exploit their spectrally-smooth nature. By Fourier-transforming our visibilities along the frequency axis, the delay domain provides a natural space in which to isolate foregrounds. Physically, a time delay can be thought of as the extra time light from a source takes to reach one antenna compared to another one (due to slightly different traveling distances). The maximum delay a source on the sky can have is if it is located on the horizon. In this case, the delay is determined by the length of the baseline. Consequently, all spectrally-smooth foreground signals show up within a maximum delay (or `horizon limit') given a baseline length. The EoR signal, however, is believed to mimic noise which has a random frequency structure and is not confined to be within the horizon limit. A simple way to avoid foregrounds without removing the cosmological signal, therefore, is to make our power spectrum measurements outside of the horizon limit that contains the bright foregrounds.

In practice, we filter our calibrated visibilities by de-convolving out our sampling function (which contains flags due to RFI) from our delay-domain visibilities using a CLEAN-like algorithm and restricting our clean components to inside the horizon limit (set by the baseline length) plus a $15 ns$ buffer. The Fourier-transformed clean components are then subtracted from our visibilities. This filtering process is performed on a per-baseline, per-integration basis, and we achieve a brightness suppression of $~2$ orders of magnitude in our visibilities.
